---
output:
  word_document: default
  html_document: default
---
Assignment Week 2: Reproducable research 
===============
*What is mean total number of steps taken per day?*

**1. Calculate the total number of steps taken per day**

```{r}
if (!file.exists("activity.csv") )
    {
     dlurl <- 'http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
     download.file(dlurl,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
     unzip('repdata%2Fdata%2Factivity.zip')
}

data<- read.csv ("activity.csv")
```


```{r}
mean_steps_day <- aggregate(steps~date, data, sum)
```

**Make a histogram of the total number of steps taken each day**
```{r}
hist(mean_steps_day$steps, main = paste("Total daily steps"), col="blue",xlab="Number of Steps")
```
**Calculate and report the mean and median of the total number of steps taken per day**
```{r}
meansteps<-mean( mean_steps_day$steps)
meansteps
mediansteps<-median(mean_steps_day$steps)
```

The mean is 10766.19 and median is 1076


*What is the eaverage daily activity pattern?*
**Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)**


```{r}
Steps_interval<-aggregate(steps~interval, data, mean)
plot(Steps_interval$interval, Steps_interval$steps, type="l",  xlab= "int", ylab= "Number of steps", main= "Av. Nunber of steps/Day")
```

**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**
  
```{r}  
max_int<- Steps_interval[which.max(Steps_interval$steps),1]
max_int
```

The interval with the maximum steps is 835


*Imputing missing values*


**Calculate and report the total number of missing values in the dataset**

```{r}
NAtot<-sum(!complete.cases(data))
NAtot
```

there are 2304 missing values in the dataset

**Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.**


```{r}
StepsAverage <- aggregate(steps ~ interval, data = data, FUN = mean)
fillNA <- numeric()
for (i in 1:nrow(data)) {
  obs <- data[i, ]
  if (is.na(obs$steps)) {
    steps <- subset(StepsAverage, interval == obs$interval)$steps
  } else {
    steps <- obs$steps
  }
  fillNA <- c(fillNA, steps)
}
StepsAverage
```


**Create a new dataset including the imputed missing values**
```{r}
new_data<- data
new_data$steps <- fillNA
```

**Make a histogram of the total number of steps taken each dat and culculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?**

```{r}
StepsTotalUnion <- aggregate(steps ~ date, data = new_data, sum, na.rm = TRUE)
hist(StepsTotalUnion$steps, main = paste("Daily step total"), col="green", xlab="Number of Steps")
#Create Histogram to show difference. 
hist(mean_steps_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps", add=T)
legend("topright", c("Imputed", "Non-imputed"), col=c("green", "blue"), lwd=10)

rmeantotal <- mean(StepsTotalUnion$steps)
rmeantotal
```


mean total is 10766.19

```{r}
rmediantotal <- median(StepsTotalUnion$steps)
rmediantotal
```

median of total is 10766.19

```{r}
mediandiff <- rmediantotal - mediansteps
mediandiff
```

there is a difference with the first part of the assignment by 1.188679

```{r}
meandiff<-rmeantotal -meansteps
meandiff
```

there is no difference in the mean however


*3d Are there differences in activity patterns between weekdays and weekends?*

**Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.**
```{r}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", 
              "Friday")
new_data$dow = as.factor(ifelse(is.element(weekdays(as.Date(new_data$date)),weekdays), "Weekday", "Weekend"))
StepsTotalUnion <- aggregate(steps ~ interval + dow, new_data, mean)
```

**Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.**

```{r}
library(lattice)
xyplot(StepsTotalUnion$steps ~ StepsTotalUnion$interval|StepsTotalUnion$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```

as seen in the plot there are difference sin activity pattrrns.  The weekend show more variance in mean steps.  While, the weekdays peak at around 800.  However, both do show similar activity patters dispite these major differneces.